---
title: "superstore_analyze"
author: "Rama"
date: "2023-04-02"
output: 
  html_document: 
    fig_width: 9
---



> About Dataset

https://www.kaggle.com/datasets/vivek468/superstore-dataset-final 

* Metadata
* Row ID => Unique ID for each row.
* Order ID => Unique Order ID for each Customer.
* Order Date => Order Date of the product.
* Ship Date => Shipping Date of the Product.
* Ship Mode=> Shipping Mode specified by the Customer.
* Customer ID => Unique ID to identify each Customer.
* Customer Name => Name of the Customer.
* Segment => The segment where the Customer belongs.
* Country => Country of residence of the Customer.
* City => City of residence of of the Customer.
* State => State of residence of the Customer.
* Postal Code => Postal Code of every Customer.
* Region => Region where the Customer belong.
* Product ID => Unique ID of the Product.
* Category => Category of the product ordered.
* Sub-Category => Sub-Category of the product ordered.
* Product Name => Name of the Product
* Sales => Sales of the Product.
* Quantity => Quantity of the Product.
* Discount => Discount provided.
* Profit => Profit/Loss incurred.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


> Prepare 

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
library(dplyr)
library(janitor)
library(tidyverse)
```


> Read Data

```{r message=FALSE, warning=FALSE, paged.print=TRUE}


superstore <- read_csv("Sample - Superstore.csv") %>% clean_names()
superstore <- superstore %>% select(-c(row_id))
```

> Changing Data type of date using lubridate

```{r message=FALSE, warning=FALSE, paged.print=TRUE}

library(lubridate)
superstore$order_date <- mdy(superstore$order_date)
superstore$ship_date <-  mdy(superstore$ship_date)
```
 
> Cleaning 
 
 
```{r message=FALSE, warning=FALSE, paged.print=TRUE}
# Checking Duplicate Value
sum(duplicated(superstore))
# Find Index Duplicate Value
which(duplicated(superstore))
# Find rows 3407 and remove it
superstore[3407,]
superstore <- superstore[-3407,]
```

> check na values


```{r message=FALSE, warning=FALSE, paged.print=TRUE}

sapply(superstore, function(x)sum(is.na(x)))
```


> Quick Summary


```{r message=FALSE, warning=FALSE, paged.print=TRUE}
summary(superstore)
```


## Task ~ Sales Analyze 

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
superstore %>%  mutate(year = year(order_date)) %>%
  group_by(year) %>% 
  summarise(sales_year = sum(sales)) %>% 
  mutate(change = (sales_year/lag(sales_year))-1, 
         change_percent = scales::percent(change)) %>% 
  select(-c(change)) %>% 
  ggplot(aes(year, sales_year))+ 
  geom_line(color = "blue", linewidth = 1)+
  geom_point(color = "blue")+
  scale_y_continuous(labels = scales::dollar)+
  geom_label(aes(label = change_percent))+
  labs(x = "Year",
       y = "Sales", 
       title = "Sales From 2014 to 2017")
```


```{r message=FALSE, warning=FALSE, paged.print=TRUE}
superstore %>% select(order_date, sales) %>% 
  mutate(month = month(order_date, label = TRUE),
         year = year(order_date)) %>%
  group_by(year, month) %>% 
  summarise(total = sum(sales)) %>% 
  ggplot(aes(month, total, group = year))+
  geom_line(color = "red", linewidth = 1)+
  geom_point()+
  facet_wrap(~year)+
  scale_y_continuous(labels = scales::dollar)+
  labs(title= "Sales From 2014-2017 by Month")
```


> Does the discount given and the season have an impact on sales?


```{r message=FALSE, warning=FALSE, paged.print=TRUE}
# adding variable season, discounted

   s_diag <-  superstore %>% select(order_date, sales, profit, quantity, discount) %>% 
  mutate(season = case_when(
    month(order_date) %in% 3:5~"Spring",
    month(order_date) %in% 6:8~"Summer",
    month(order_date) %in% 9:11~"Fall",
    TRUE ~"Winter"),
    holiday = ifelse(weekdays(order_date)%in%c("Saturday", "Sunday"), "Weekend", "Weekday"),
    discounted = ifelse(discount>0, "Yes", "No"))
head(s_diag)
```



```{r message=FALSE, warning=FALSE, paged.print=TRUE}
 s_diag %>%
   mutate(month = month(order_date, label = TRUE),
          year = year(order_date)) %>%
   group_by(year, month, season) %>%
   summarise (total = sum(sales),
              discounted = sum(discounted == "Yes")) %>%
   ggplot(aes(month, total, fill = season))+
   geom_bar(stat='identity')+
   facet_wrap(~year)+
   geom_text(aes(label = discounted), position = position_stack(vjust = 0.5))+
   scale_y_continuous(labels = scales::dollar)+
   theme(legend.position = "bottom")+
   labs(title = "Total Sales in each year and total discount given")
```

> discounts do not affect sales value, but after entering the end of the year during the fall season customers tend to make purchases at the beginning, late fall, and end of the year (early winter). the total discount given is also more than other seasons. in early spring, sales are always increasing in every year from the previous month


```{r message=FALSE, warning=FALSE, paged.print=TRUE}

 s_diag %>%
   mutate(month = month(order_date, label = TRUE)) %>%
   group_by(month, holiday, season) %>%
   summarise (avg_sales = mean(sales)) %>%
   ggplot(aes(month, avg_sales, fill = holiday)) +
   geom_bar(stat='identity', position = position_dodge())+
   scale_y_continuous(labels = scales::dollar)+
   labs(title = "Avg_sales and type of days")+
   theme(legend.position = "bottom")
 
   
```

> From the graph, we know that holidays didn't affect sales. Companies don't need to make a decision about whether to provide discounts on weekdays or weekends. However, for the 4 months (mar, sep, nov, dec), we may consider providing a discount or promotion, either on weekdays or weekends.


## Task ~ Analyze Customer Segments


```{r message=FALSE, warning=FALSE, paged.print=TRUE}
superstore %>% 
  group_by(segment) %>% 
  summarise(total_sales = sum(sales),
            total_profit = sum(profit),
            total_orders = n()) %>% 
  mutate_at(vars(total_sales, total_profit), scales::dollar)
```
```{r message=FALSE, warning=FALSE, paged.print=TRUE}
superstore %>% 
  mutate(year = year(order_date)) %>% 
  group_by(year, segment) %>% 
  summarise(total_sales = sum(sales), group = 1) %>% 
  ggplot(aes(year, total_sales, fill = segment))+ 
  geom_bar(stat = 'identity')+
  scale_y_continuous(labels = scales::dollar)+
  theme_minimal()+
  facet_wrap(~segment)+
  theme(legend.position = "none")
```

> From the table, we know that Segment of consumer provide the most sales and profit



> Region and State

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
#Checking Unique Region 

unique(superstore$region)

#Checking Unique State

unique(superstore$state)
```


> Which region the most sales?


```{r message=FALSE, warning=FALSE, paged.print=TRUE}

superstore %>% group_by(region) %>% 
  summarise(total_sales = sum(sales)) %>% 
  arrange(desc(total_sales)) %>% 
  mutate_at(vars(total_sales), scales::dollar)
```

> TOP 3 Total Sales In State By Region

```{r}
superstore %>% 
group_by(region,state) %>% 
  summarise(total_sales = sum(sales)) %>% 
  arrange(desc(total_sales)) %>% 
  slice(1:3) %>%
  ggplot(aes(reorder(state, -total_sales), total_sales, fill = region))+
  geom_bar(stat = 'identity')+
  labs(title = 'Top 3 Sales In State By Region',
       x = 'state')+
  scale_y_continuous(labels = scales::dollar)+
  facet_wrap(~region, scales = "free_x")+
  theme(legend.position = "none")
```


> TOP 10 Sales by City

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
superstore %>% 
  group_by(city) %>% 
  summarise(total_sales = sum(sales)) %>% 
  arrange(desc(total_sales)) %>%
  slice(1:10) %>% 
  ggplot(aes(reorder(city, total_sales), total_sales, fill = 'red'))+
  geom_bar(stat = 'identity')+
  coord_flip()+
  scale_y_continuous(labels = scales::dollar)+
  theme(legend.position = "none")+
  labs(x = "City")
```


> How many products are purchased in each segment?


```{r message=FALSE, warning=FALSE, paged.print=TRUE}

superstore %>% mutate(year = year(order_date)) %>% 
  group_by(year, segment) %>% 
  summarise(total_qty = sum(quantity)) %>% 
  ggplot(aes(year, total_qty, fill = segment))+
  geom_bar(stat = 'identity', position = position_dodge())+
  labs(x = "")

```

## Task ~ Produk Analyze



```{r message=FALSE, warning=FALSE, paged.print=TRUE}
p_quantity <- 
  superstore %>% 
  group_by (product_name) %>% 
  summarise (total_quantity = sum(quantity)) %>% 
  arrange(desc(total_quantity)) %>% 
  slice(1:10) %>% 
  mutate(product_name = str_replace(product_name, "(^[^ ]* [^ ]* [^ ]*).*", "\\1"),
         product_name = str_wrap(product_name, width = 15)) %>% 
  # "(^[^ ]* [^ ]* [^ ]*).*" menunjukkan pola tiga kata pertama (di mana ^ menunjukkan awal string, 
  # [^ ]* menunjukkan setiap karakter kecuali spasi, dan [ ] menunjukkan spasi)
  # diikuti oleh karakter apa pun (.*). "\\1" menunjukkan pengembalian dari tiga kata pertama.
  # str_wrap digunakan untuk membungkus kata dengan lebar yang telah ditentukan
  ggplot(aes(reorder(product_name,total_quantity), total_quantity, fill=total_quantity))+
  geom_bar(stat = 'identity')+
  coord_flip()+
  scale_fill_gradient2(mid="purple", high = "black")+
  labs(x = "Product Name",
       y = "Total_Quantity")+
  theme_minimal()+theme(legend.position = "none")


p_sales <- 
  superstore  %>% 
  group_by(product_name) %>% 
  summarise (total_sales = sum(sales)) %>% 
  arrange(desc(total_sales)) %>% 
  slice(1:10) %>% 
  mutate(product_name = str_replace(product_name, "(^[^ ]* [^ ]* [^ ]*).*", "\\1"),
         product_name = str_wrap(product_name, width = 15)) %>%
  ggplot(aes(reorder(product_name, total_sales), total_sales, fill = total_sales))+
  geom_bar(stat = 'identity')+
  coord_flip()+
  scale_fill_gradient2(mid = "purple", high = "black")+
  labs(x = "Product Name",
       y = "Total_Sales")+
  theme_minimal()+theme(legend.position = "none")+
  scale_y_continuous(labels = scales::dollar)


library(grid)
library(gridExtra)
grid.arrange(p_quantity, p_sales, ncol=2)


```

> charts showing top 10 product based on quantity and sales



> Top 10 Product by Loss and Profitable

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
superstore %>% select(product_name, sales, profit, discount) %>% 
  mutate(product_name = str_replace(product_name, "(^[^ ]* [^ ]* [^ ]*).*", "\\1"))%>% 
  arrange(profit) %>% 
  slice(1:10) %>% 
  mutate(sales = scales::dollar(sales),
         profit = scales::dollar(profit),
         discount = paste0(round(discount*100), "%"))
         

superstore %>% select(product_name, sales, profit, discount) %>% 
  mutate(product_name = str_replace(product_name, "(^[^ ]* [^ ]* [^ ]*).*", "\\1"))%>% 
  arrange(desc(profit)) %>% 
  slice(1:10) %>% 
  mutate(sales = scales::dollar(sales),
         profit = scales::dollar(profit),
         discount = paste0(round(discount*100), "%"))
         
```


